---
title: "BIOL432_Assignment3_Markdown"
author: "Amanda Zacharias"
date: "1/26/2021"
output: html_document
---

# Project Info

**Date**: January 26th, 2021
**GitHub Repository Link**: [link](https://github.com/ForgetfulCow/BIOL432_Assignment3)
**Username**: ForgetfulCow

# Load the packages
```{r}
library(dplyr)
library(knitr)
```

# Load Data
Reads in the dataset. Makes a data.frame object called 'Fallo'.
```{r}
Fallo <- read.csv("InputData/FallopiaData.csv")
```

# Preview the Data
Outputs the first 6 rows of the Fallo data.frame object.
```{r}
head(Fallo)
```

> Note: Each time dataset is manipulated, a new dataset is created. This is to preserve the data and its changes.

# Remove rows with 'Total' biomass < 60.
Using `filter()` function to subset the Fallo dataset and only include rows where 'Total' biomass is greater or equal to 60.

\n

Using `head()` function to output the first 6 rows of the new Fallo1 dataset.
```{r}
Fallo1 <- filter(Fallo, Total >= 60)
head(Fallo1)
```

# Reorder the columns so that they are in order: 'Total', 'Taxon', 'Scenario', 'Nutrients', and remove the other columns
Using `transmute()` function, manipulate the inclusion and order of variables in the Fallo1 dataset. 

\n

Using `head()` function to output the first 6 rows of the new Fallo1 dataset.
```{r}
Fallo2 <- transmute(Fallo1, Total, Taxon, Scenario, Nutrients)
head(Fallo2)
```

# Make a new column TotalG, which converts the 'Total' column from mg to grams AND replaces Total with TotalG.
Using `mutate()` to create the new column with 'Total' values divided by 1,000, thus converting miligrams to grams. 

\n

Using `rename()` to change the name, 'Total', to 'TotalG'
Using `head()` to print the first 6 rows of the new Fallo3 dataset.
```{r}
Fallo3 <- Fallo2 %>% 
  mutate(Total = Total/1000) %>% 
  rename(TotalG = Total)
head(Fallo3)
```

# Write a custom function that will take two inputs from the user: 1. a vector of data to process (e.g. column from a data.frame object) and 2. a string that defines what calculations to perform.
Inputs: 

+ "vec" = data to process, object type: vector
+ "calc" = calculation to perform, object type: string

Possible calculations: 

+ "Average" = calculate the average value
+ "Sum" = calculate the sum of values
+ "Observations" = count the number of observed values
+ If any other string is input, output an error to the user

Output: 

+ Calculation (numerical) or error message
```{r}
maths_func <- function(vec=Default, calc=Default){
  if (calc == "Average"){
    Average = mean(vec)
    return(Average)
  } # ends Average statement
  if (calc == "Sum"){
    Sum = sum(vec)
    return(Sum)
  } # ends if Sum statement
  if (calc == "Observations"){
    Length = length(vec)
    return(Length)
  } # ends if Observation statement
  else {
    msg <- 'Input is invalid. Please enter, "Average", "Sum", or "Observations"'
    print(paste(msg))
  } # ends else statement
} # ends function
```

# Write some R code that uses your function to count the total number of observations in the 'Taxon' column.
Calls the custom maths_func function to do calculations. Stores the function's output in the obs_Taxon variable.

\n

Uses `print()` to display the output to the user.
```{r}
obs_Taxon <- maths_func(Fallo3$Taxon, "Observations")
print(obs_Taxon)
```

# Write some R code that uses your function to calculate the average TotalG for each of the two nutrient concentrations
Calls the custom maths_func function to do calculations. Stores the function's output in the nutrient_grp variable.

\n

Uses `print()` to display the output to the user.

\n

The `summarize()` function will output a warning message that there is grouping. This is to be expected and there is no error.

\n

The `kable()` function is used to make the output of `summarize()` more attractive. 
```{r}
nutrient_grp <- Fallo3 %>% 
  group_by(Nutrients) 
ave_table <- summarize(nutrient_grp, ave_tg = maths_func(nutrient_grp$TotalG, "Average"))
kable(ave_table, caption = "Average by Nutrient Level Table")
```


