---
title: "BIOL432_Assignment4_Markdown"
author: "Amanda Zacharias"
date: "2/2/2021"
output: html_document
---

# Project Info

**Date**: February 2nd, 2021
**GitHub Repository Link**: [link](https://github.com/ForgetfulCow/BIOL432_Assignments)
**Username**: ForgetfulCow

# Load the packages

package loaded: dplyr

```{r}
library(dplyr)
```

# Import Data

Reading in the measurements.csv dataset from the InputData folder of this project. Dataset is saved in the "MData" object. `head()` is used to view the first 6 rows of the dataset.

```{r}
MData <- read.csv("InputData/measurements.csv")
head(MData)
```

# Add a new column called Sp that shortens the genus name in the Organism column to only the first letter and period. 

## Example: *Homo sapiens* becomes *H. sapiens* in the Sp column.

## Use dplyr and regular expression commands

`mutate()` is used to add the new column, Sp, to the dataset. For the purposes of preserving data, a new dataset is created with the extra column. 

`sub()` is used to substite the first word of an organism's species name with its first leter and a period symbol. Example: John Smith --> J. Smith.

```{r}
MData2 <- MData %>% 
  mutate(Sp = sub("(\\w)\\w+","\\1.",Organism))

head(MData2)
```

# Create a new dataset but with Sp column instead of the Organism column, moved to the beginning in place of the original Organism column. 

`transmute()` is used to reorder the MData2 dataset and replace the Organism column with the Sp column. The Organism column is removed.

\n

Alterations to MData2 are saved in the MData3 object.

```{r}
MData3 <- MData2 %>% 
  transmute(Sp, Limb.width, Limb.length, Observer)

head(MData3)
```

# Save the dataset as a new file called MeasShort.csv

Template for saving output/data is taken from [Colautti's R Crash Course](https://colauttilab.github.io/RCrashCourse/1_fundamentals.html#6_save_output)

`write.csv()` will write the MData3 data.frame object to a .csv file called "MeasShort.csv". File is saved in the Output file of this project.

```{r}
write.csv(MData3, "Output/MeasShort.csv")
```
